CF Registry Analysis

### 2024-11-04 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
minor updates - going to try just running three separate models per Kristen

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical")
        ) %>%
        add_p() %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  ) %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(screening, aes(
  x = ReviewYear, y = value,
  color = age_group
)) +
  geom_line() +
  xlab("Year") +
  facet_wrap(~name, nrow = 3) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group") +
  theme(text = element_text(size = 20), axis.title.y = element_blank())
screening_plot
ggsave("./Dissemination/Figures/screening_rates.png",
  plot = screening_plot, width = 12, height = 9, units = "in"
)
```

@@

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(time = as.numeric(tstart), time2 = as.numeric(tstop), event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression()
ggcoef_table(cox_res, show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

@@

```{r results='asis'}
f_vx770 <- update(f, . ~ . + A_Vx770)
vx770_cox_res <- coxph(f_vx770,
  data = annual[annual$ReviewYear >= 2011, ],
  cluster = eDWID
)
vx770_cox_res %>%
  tbl_regression()
ggcoef_table(vx770_cox_res,
  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
)
```

@@

```{r}
f_vx809comb <- update(f, . ~ . + A_VX809comb)
vx809comb_cox_res <- coxph(f_vx809comb,
  data = annual[annual$ReviewYear >= 2015, ],
  cluster = eDWID
)
vx809comb_cox_res %>%
  tbl_regression()
ggcoef_table(vx809comb_cox_res,
  show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
)
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression() %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_table(cox_res_split,
    show_p_values = F, add_reference_rows = T, signif_stars = F, colour = F
  ) %>%
    print()
  cat("\n")
  cat("\n")
}))
```

>>>>

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/shared-resources/Visualization/forest_plots.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_perc_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical"),
          statistic = list(c(
            "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
            "A_FVC_pct_predicted"
          ) ~ "{mean} ({sd})")
        ) %>%
        add_p(test = list(c(
          "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
          "A_FVC_pct_predicted"
        ) ~ "t.test")) %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  )
kable(
  screening %>%
    select(ReviewYear, age_group, `Screened OGTT`) %>%
    mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
    pivot_wider(names_from = age_group, values_from = `Screened OGTT`),
  caption = "% Screened by OGTT"
)
screening <- screening %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(screening, aes(
  x = ReviewYear, y = value,
  color = age_group
)) +
  geom_line() +
  xlab("Year") +
  facet_wrap(~name, nrow = 3) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group") +
  theme(text = element_text(size = 20), axis.title.y = element_blank())
screening_plot
ggsave("./Dissemination/Figures/screening_rates.png",
  plot = screening_plot, width = 12, height = 9, units = "in"
)
```

@@

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(time = as.numeric(tstart), time2 = as.numeric(tstop), event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

@@

```{r results='asis'}
f_vx770 <- update(f, . ~ . + A_Vx770)
vx770_cox_res <- coxph(f_vx770,
  data = annual[annual$ReviewYear >= 2011, ],
  cluster = eDWID
)
vx770_cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(vx770_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
f_vx809comb <- update(f, . ~ . + A_VX809comb)
vx809comb_cox_res <- coxph(f_vx809comb,
  data = annual[annual$ReviewYear >= 2015, ],
  cluster = eDWID
)
vx809comb_cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(vx809comb_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r results='asis'}
invisible(lapply(time_varying, function(v) {
  # Formulas
  new_f <- as.formula(paste0(
    y, "~", paste0(xvars[-which(xvars == v)], collapse = "+"),
    paste0("+", v, "*strata(tgroup)")
  ))
  new_f_emm <- as.formula(paste0(
    "~", paste0(v, "|tgroup")
  ))
  # This could be done more efficiently but it seems to work
  # Fit one stratified model and one with a regular interaction term
  cox_res_split <- coxph(new_f, data = annual_split, cluster = eDWID)
  # Use the regular interaction term to get means for each time period
  cat("\n")
  cat("\n")
  cat("####", v)
  cat("\n")
  cat("\n")
  cox_res_split %>%
    tbl_regression(exponentiate = T) %>%
    print()
  cat("\n")
  cat("\n")
  emm <- emmeans(cox_res_split, new_f_emm, nuisance = xvars[-which(xvars == v)])
  emm <- data.frame(emm)
  # Plot base model (without time varying)
  ggcoef_table(cox_res_split,
    show_p_values = F, add_reference_rows = T, exponentiate = T,
    signif_stars = F, colour = F
  ) %>%
    print()
  cat("\n")
  cat("\n")
}))
```
